---
title: "Sobol Assignment - Wind Speed Sensitivity"
theme: solarized
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(sensitivity)
library(tidyverse)
library(purrr)
library(here)
```

## Assignment Part I {.scrollable}

Choose one of the 3 papers below that provide an example of sensitivity analysis of model parameters. After going through the paper, write a paragraph describing how results of the *sensitivity analysis* reported on in the paper might contribute to understanding (or prediction) within an environmental problem solving or management context.

[Snow modeling](Snow_sobol.pdf)

The results of the Snow Modeling paper indicate how different models perform better under different circumstances, and how calibrating a model, which is apparently not a common practice, can improve outputs. One of the models, the SNOW-17 model performs better in a shallow snowpack, whereas the VIC model performs better with deeper snowpacks. Additionally, the SNOW-17 model is in general more stable, but responds less to dynamic parameters. This makes is a less-than-ideal choice for climate modeling. In contrast, the VIC model contains more variables, and is more sensitive to a range of values, including albedo, meaning it is better suited to model unpredictable and varaible condition. This also means that careful calibration is required for optimum model performance.

## Assignment Part 2 {.scrollable}

Recall our model of atmospheric conductance $$ C_{at} = \frac{v_m}{6.25*{ln(\frac{z_m-z_d}{z_0})}^2} $$ $$ z_d = k_d*h $$ $$ z_0 = k_0*h $$

$z_m$ is the height at which windspeed is measured - must be higher than the vegetation (cm), it is usually measured 200 cm above the vegetation

$h$ is vegetation height (cm)

$v$ is windspeed (cm/s)

Typical values if $k_d$ and $k_o$ are 0.7 and 0.1 respectively (so use those as defaults)

## Your task {.scrollable}

Repeat the sensitivity analysis that we have been working on in class BUT lets assume that we are in a different locations - where windpeeds are substantially higher and more variable AND vegetation is shorter - See details below

Consider the sensitivity of your estimate to uncertainty in the following parameters and inputs

-   $height$

-   $k_d$

-   $k_0$

-   $v$

Windspeeds $v$ are normally distributed with a mean of 300 cm/s with a standard deviation of 50 cm/s

For vegetation height assume that height is somewhere between 3.5 and 5.5 m (but any value in that range is equally likely)

For the $k_d$ and $k_0$ parameters you can assume that they are normally distributed with standard deviation of 1% of their default values

a)  Use the Sobel approach to generate parameter values for the 4 parameters
b)  Run the atmospheric conductance model for these parameters
c)  Plot conductance estimates in a way that accounts for parameter uncertainty
d)  Plot conductance estimates against windspeed use the parameter that is 2nd in terms of total effect on response
e)  Estimate the Sobel Indices for your outut
f)  Comment on what this tells you about how atmospheric conductance and its sensitivity to variation in windspeed differs in this setting as compared to the setting that we examined in class where windspeed was lower and less variable and vegetation was taller.

Submit the Quarto on Canvas as usual

## A) Generate Parameter Values for the 4 Parameters {.scrollable}

First, generate parameter sets for the Sobol analysis

-   2 sets of samples, generate by sampling from expected distributions using sobolSalt

```{r}
source(here("Catm.R"))

# generate two examples of random number from parameter distributions

np <- 1000
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.1, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.1, n = np)
v <- rnorm(mean = 300, sd = 50, n = np)
height <- runif(min = 3.5, max = 5.5, n = np)

X1 <- cbind.data.frame(k_o, k_d, v, height = height)

# repeat sampling
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.1, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.1, n = np)
v <- rnorm(mean = 300, sd = 50, n = np)
height <- runif(min = 3.5, max = 5.5, n = np)

X2 <- cbind.data.frame(k_o, k_d, v, height = height)

# there are different versions of sobol functions that have different approaches for estimating parameters and indices, we use an approach implemented by jansen

sens_Catm_Sobol <- sobolSalt(model = NULL, X1, X2, nboot = 100)

head(X1)
head(X2)

# Take a look at the Sobol generated sensitivity object
# your parameters sets for sensitivity analysis are in X
```

## B) Run the atmospheric conductance model for generated parameters and compute Sobol Indices {.scrollable}

run model for Sobol generated parameter sets and compute indices

```{r sobolrun}
# run model for all parameter sets
# make sure you give the parameters names

parms <- as.data.frame(sens_Catm_Sobol$X)
colnames(parms) <- colnames(X1)
res <- pmap_dbl(parms, Catm)


sens_Catm_Sobol <- sensitivity::tell(sens_Catm_Sobol, res, res.names = "ga")

# main effect:  partitions variance (main effect without co-variance) - sums approximately to one
sens_Catm_Sobol$S
# useful to add names
row.names(sens_Catm_Sobol$S) <- colnames(parms)
sens_Catm_Sobol$S

# total effect - accounts for parameter interactions
row.names(sens_Catm_Sobol$T) <- colnames(parms)
sens_Catm_Sobol$T

# Both the main effect and total effect can tell us something about how the parameter influences results


print(sens_Catm_Sobol)
```

## C) Plot Conductance Estimates {.scrollable}

```{r sobolplot}
# graph two most sensitive parameters
both <- cbind.data.frame(parms, gs = sens_Catm_Sobol$y)

# look at overall gs sensitvity to uncertainty
ggplot(both, aes(x = gs)) +
  geom_histogram() +
  geom_vline(xintercept = mean(both$gs), col = "cyan")

# look at response of conductance to wind speed and roughness
#roughness is second in terms of effect on response
ggplot(both, aes(v, gs, col = k_o)) +
  geom_point() +
  labs(y = "Conductance (mm/s)", x = "Windspeed")


```

## Second order indices {.scrollable}

```{r, alternative}
sens_Catm_Sobol2 <- sobolSalt(model = NULL, X1, X2, nboot = 100, scheme = "B")

parms <- as.data.frame(sens_Catm_Sobol2$X)
colnames(parms) <- colnames(X1)
res <- pmap_dbl(parms, Catm)


sens_Catm_Sobol2 <- sensitivity::tell(sens_Catm_Sobol2, res, res.names = "ga")

# main effect:  partitions variance (main effect without co-variance) - sums approximately to one
row.names(sens_Catm_Sobol2$S) <- colnames(parms)
sens_Catm_Sobol2$S

# total effect - accounts for parameter interactions
row.names(sens_Catm_Sobol2$T) <- colnames(parms)
sens_Catm_Sobol2$T

# second order parameters interaction in controlling sensitivity
# parameters are in order, interactiosn are small here
sens_Catm_Sobol2$S2
```

## F) Atmospheric Conductance Analysis {.scrollable}

Comment on what this tells you about how atmospheric conductance and its sensitivity to variation in windspeed differs in this setting as compared to the setting that we examined in class where windspeed was lower and less variable and vegetation was taller.

Atmospheric conductance is most sensitive to windspeed, and second most sensitive to plant roughness. Increases in both correlate to an increase in conductance, and roughness is a better determinant of conductance than height. In class, with lower and less variable windspeeds, k_d was the biggest determinant of conductance. Second order interactions are farily limited in this scenario.
